---
title: "Ranking NBA Teams"
output: github_document
author: 'Brandon Huang'
---

### Table of Contents
* Barplots of Salary, Points, Experience with Average Vertical Line 

---

``` {r library, include = FALSE}
library(ggplot2)
library(dplyr) #load pacakge: dplyr 
``` 

### Creating `teams` DataFrame 

---

``` {r importing data}
setwd("/Users/brandon/Desktop/stat133/stat133-hws-fall17/hw03/data") 
teams <- read.csv(file = 'nba2017-teams.csv', stringsAsFactors = FALSE)
```

``` {r arrange teams salary in desc order, include = FALSE}
teams <- arrange(teams, desc(salary)) 
``` 

---

## Barplots of Salary, Points, Experience with Average Vertical Line 

---

#### 1. Barplot of NBA Teams Ranked by Total Salary

``` {r horizontal barplot ranked by salary with vertical mean salary line}
ggplot(teams, aes(x = reorder(team, salary), y = salary)) +
  geom_bar(stat = 'identity', fill = 'grey', size = 1) +
  coord_flip() + 
  labs(title = 'NBA Teams ranked by Total Salary') +
  xlab(label = 'Teams') +
  ylab(label = 'Salary (in millions)') +
  geom_hline(yintercept = mean(teams$salary), colour = 'red', size = 2)
```   

---

#### 2. Barplot of NBA Teams Ranked by Total Points

``` {r horizontal barplot ranked by total points with vertical mean points line} 
ggplot(teams, aes(x = reorder(team, points), y = points)) +
         geom_bar(stat = 'identity', fill = 'grey', size = 1) +
         coord_flip() + 
         xlab(label = 'Teams') +
         ylab(label = 'Points') + 
         geom_hline(yintercept = mean(teams$points), colour = 'red',
                    size = 2) + 
  labs(title = 'NBA Teams Ranked by Points') 
``` 

---

#### 3. Barplot of NBA Teams Ranked by Total Efficiency

``` {r horizontal barplot ranked by total efficiency with mean efficiency}
ggplot(teams, aes(x = reorder(team, efficiency), y = efficiency)) + 
  geom_bar(stat = 'identity', fill = 'grey', size = 1) + 
  xlab(label = 'Teams') +
  ylab(label = 'Efficiency') + 
  labs(title = 'NBA Teams Ranked by Efficiency') +
  geom_hline(yintercept = mean(teams$efficiency), colour = 'red', size = 2) +
  coord_flip()
``` 

---

##### Common Errors

---
* geom_hline:
  * forgetting to include "y intercept"
  * "colour" not "color" 
* using plural versions of parameters 
* coord_flip not flip_coord 
* because of coord_flip:
  * using x as y, and y as x 
  * using vline instead of hline
  * xlab and ylab mix up 

---

### PCA 

---
``` {r graph of PC1 and PC2}
teams <- teams[-1]
teams1 <- data.frame(teams[ , c(1, 4, 5, 6, 8, 9, 10, 11, 12, 13, 14)]) #data frame of variables # needed
pca <- prcomp(teams1[-1], scale. = TRUE) #pca of teams1
composite <- data.frame(team = teams[ , 1],pc1 = pca$x[ , 1], pc2 = pca$x[ , 2], stringsAsFactors = FALSE) #dataframe of teams, pc1, pc2
# composite$pc1 <- -1 * composite$pc1 #pc1 times -1 ... still don't know why 
composite1 <- data.frame(team = teams[ , 1], pc1 = pca$x[ , 1], pc2 = pca$x[ , 2], stringsAsFactors = FALSE) #same data frame as composite but for the next section
# composite1$pc1 <- -1 * composite1$pc1 #same -1 conversion 
ggplot(composite, aes(pc1, pc2)) + geom_point() + geom_text(aes(label = teams$team)) + labs(title = "PCA plot (PC1 and PC2)") + geom_hline(yintercept = 0) + geom_vline(xintercept = 0)  #scatterplot of pca 1 and pca 2
``` 

---

### Index Based on PC1

---
``` {r scale pc1 and barplot}
composite1$pc1 <- (composite1$pc1 - min(composite1$pc1))/(max(composite1$pc1) - min(composite1$pc1))

ggplot(composite1, aes(x = reorder(team, pc1), y = pc1)) +
         geom_bar(stat = 'identity', fill = 'grey', size = 1) +
         coord_flip() + 
         xlab(label = 'Team') +
         ylab(label = 'First PC(scaled from 0 to 100)') + 
  labs(title = 'NBA Teams ranked by scaled pc1') 
```

Interesting how Cleveland is ranked so low. 

--- 

##### Common Errors

---

* I NEED TO READ THE INSTRUCTIONS AND PROBLEMS MORE CAREFULLY 
