---
title: "lab05_brandon_huang"
output: html_document
author: 'Brandon Huang'
---

### Table of Contents
> * Get started with `'dplyr'`


``` {r include = FALSE}
library(dplyr)
library(ggplot2)
```

``` {r}
dat <- read.csv("/Users/brandon/Desktop/stat133/stat133-hws-fall17/hw01/nba2017-players.csv", stringsAsFactors = FALSE)
```

##### Hadley's list of verbs

* `filter()`: select rows by condition, `slice()`: select rows by position, `select()`: select columns by names [subsetting and selecting rows and columns]
* `mutate()`: add new variables
* `arrange()`: reorder rows
* `summarise()`: reduce variables to values
* `group_by()`: grouped (aggregate) operations 

###Your Turn: 
* use slice() to subset the data by selecting the first 5 rows.

``` {r}
slice(dat, 1:5)
```

* use slice() to subset the data by selecting rows 10, 15, 20, ..., 50.
``` {r}
slice(dat,seq(10, 50, 5))
``` 

* use slice() to subset the data by selecting the last 5 rows.
``` {r}
slice(dat, nrow(dat)-5:nrow(dat))
``` 

* use filter() to subset those players with height less than 70 inches tall.
``` {r}
filter(dat, height < 70)
```

* use filter() to subset rows of Golden State Warriors ('GSW').
``` {r}
filter(dat, team == 'GSW')
```

* use filter() to subset rows of GSW centers ('C').
``` {r}
filter(dat, team == 'GSW' & position == 'C') 
```

* use filter() and then select(), to subset rows of lakers ('LAL'), and then display their names.
``` {r}
select(filter(dat, team == 'LAL'), player)
``` 

* use filter() and then select(), to display the name and salary, of GSW point guards
``` {r}
select(filter(dat, team == 'GSW'), c(player, salary))
```

* find how to select the name, age, and team, of players with more than 10 years of experience, making 10 million dollars or less.
``` {r}
select(filter(dat, experience > 10 & salary < 10000000), c(player, age, team))
```

* find how to select the name, team, height, and weight, of rookie players, 20 years old, displaying only the first five occurrences (i.e. rows)
``` {r}
slice(select(filter(dat, experience == 0), c(player, team, height, weight)), 1:5)
```

### Adding New Variables and Reordering Rows: `mutate()`, `arrange()` 

``` {r}
gsw <- filter(dat, team == 'GSW') 
```

decreasing: arrange(dataset, desc(column)) 

Your Turn: 
* using the data frame gsw, add a new variable product with the product of height and weight.

``` {r}
gsw <- mutate(gsw, product = height * weight)
```

* create a new data frame gsw3, by adding columns log_height and log_weight with the log transformations of height and weight.
``` {r}
gsw3 <- mutate(gsw, log_height = log(height), log_weight = log(weight))
```

* use the original data frame to filter() and arrange() those players with height less than 71 inches tall, in increasing order.
``` {r}
arrange(filter(dat, height < 71), height) 
```

* display the name, team, and salary, of the top-5 highest paid players
``` {r} 
select(slice(arrange(dat, desc(salary)), 1:5), c(player, team, salary))
```

* display the name, team, and points3, of the top 10 three-point players
``` {r}
select(slice(arrange(dat, desc(points3)), 1:10), c(player, team, points3))
```

* create a data frame gsw_mpg of GSW players, that contains variables for player name, experience, and min_per_game (minutes per game), sorted by min_per_game (in descending order)
``` {r}
gsw <- filter(dat, team == 'GSW') 
gsw_mpg <- arrange(select(mutate(gsw, min_per_game = minutes/games), c(player, experience, min_per_game)), desc(min_per_game))
```

### `summarize()` and `group_by()`

Your Turn: 
* use summarise() to get the largest height value.
``` {r}
summarise(dat, max_height = max(height)) 
``` 

* use summarise() to get the standard deviation of points3
``` {r}
summarise(dat, sd(points3))
``` 

* use summarise() and group_by() to display the median of three-points, by team.
``` {r}
summarise(group_by(dat, team), median(points3))
```

* display the average triple points by team, in ascending order, of the bottom-5 teams (worst 3pointer teams)
``` {r}
slice(arrange(summarise(group_by(dat, team), mean_points3 = mean(points3)), mean_points3), 1:5)
```

* obtain the mean and standard deviation of age, for Power Forwards, with 5 and 10 years (including) years of experience.
``` {r}
summarise(filter(dat, experience == c(5:10) & position == 'PF'), sd_age = sd(age), mean_age = mean(age)) 
``` 

### GGPLOT

```{r}
#scatterplot option 1 
ggplot(dat) + geom_point(aes(x = points, y = salary))

#scatterplot option 2 
ggplot(dat, aes(x = points, y = salary)) + geom_point() 

#distinguish position by color
ggplot(dat, aes(x = points, y = salary)) + geom_point(aes(color = position)) 

#modify size of dots accordingly
ggplot(dat, aes(x = points, y = salary)) + geom_point(aes(color = position, size = points3))  

#transparency with alpha, alpha is outside of aes because it is not using columns 
ggplot(dat, aes(x = points, y = salary)) + geom_point(aes(color = position, size = points3), alpha = 0.7)
```

#### MY TURN

* Use the data frame gsw to make a scatterplot of height and weight
``` {r}
ggplot(dat, aes(x = height, y = weight)) + 
  geom_point() 
```

* Find out how to make another scatterplot of height and weight, using geom_text() to display the names of the players
``` {r}
ggplot(dat, aes(x = height, y = weight)) + 
  geom_point() + 
  geom_text(aes(label = player))
```

* Get a scatter plot of height and weight, for ALL the warriors, displaying their names with geom_label()
``` {r}
ggplot(gsw, aes(x = height, y = weight)) + 
  geom_point() + 
  geom_text(aes(label = player))
```

* Get a density plot of salary (for all NBA players)
``` {r}
ggplot(dat, aes(x = salary)) + 
  geom_density()
```

* Get a histogram of points2 with binwidth of 50 (for all NBA players)
``` {r}
ggplot(dat, aes(x = points2)) + 
  geom_histogram(binwidth = 50)
```

* Get a barchart of the position frequencies (for all NBA players)
``` {r}
position = dat$position

ggplot(dat) + 
  geom_bar(aes(position))
```

* Make a scatterplot of experience and salary of all centers, and use geom_smooth() to add a regression line
``` {r}
ggplot(dat, aes(x = dat$experience, y = dat$salary)) + 
  geom_smooth(method = lm)
```

* Repeat the same scatterplot of experience and salary of all centers, but now use geom_smooth() to add a loess line
``` {r}
ggplot(dat, aes(x = experience, y =salary)) + 
  geom_smooth(method = loess)
``` 

### Faceting

`facetwrap`
``` {r}
#scatterplot by position
ggplot(data = dat, aes(x = points, y = salary)) + 
  geom_point() + 
  facet_wrap(~position) 
``` 

`facetgrid`
``` {r}
# scatterplot by position
ggplot(data = dat, aes(x = points, y = salary)) + 
  geom_point(aes(color = position), alpha = 0.7) + 
  facet_grid(~ position) + 
  geom_smooth(method = loess)
``` 

`facetgrid horizontal` 
```{r}
ggplot(data = dat, aes(x = points, y = salary)) +
  geom_point(aes(color = position), alpha = 0.7) +
  facet_grid(position ~ .) +
  geom_smooth(method = loess)
``` 

#### My Turn 

* Make scatterplots of `experience` and `salary` faceting by `position`
```{r}
ggplot(dat, aes(x = experience, y = salary)) + 
  geom_point() + 
  facet_wrap(~position)
```

* Make scatterplots of `experience` and `salary` faceting by `team`
``` {r}
ggplot(dat, aes(x = experience, y = salary)) +
  geom_point() + 
  facet_wrap(~team) 
``` 

* Make density plots of `age` faceting by `team`
``` {r}
ggplot(dat, aes(x = age)) +
  geom_density() + 
  facet_wrap(~team) 
``` 

* Make scatterplots of `height` and `weight` faceting by `position`
``` {r}
ggplot(dat, aes(x = height, y = weight)) +
  geom_point() + 
  facet_wrap(~position)
``` 

* Make scatterplots of `height` and `weight`, with a 2-dimensional density, `geom_density2d()`, faceting by `position`
``` {r}
ggplot(dat, aes(x = height, y = weight)) +
  geom_density2d() + 
  facet_wrap(~position)
```