---
title: "hw02"
output: github_document
author: 'Brandon Huang' 
---

``` {r include = FALSE}
library(readr)
#setwd("/Users/brandon/Desktop/stat133/stat133-hws-fall17/hw02/data")
#getwd()
```

### 2. Importing the Data
``` {r}
colClasses = c(
  "character",
  "character",
  "factor", 
  "character",
  "real",
  rep("integer", 19)
)
```

``` {r}
col_brandon = list(
  col_character(),
  col_character(),
  col_factor(c('C', 'PF', 'PG', 'SF', 'SG')), 
  col_character(),
  col_double(),
  rep(col_integer(), 19)
  )
```

``` {r} 
#read_csv - readr
dat2 <- read_csv("/Users/brandon/Desktop/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv", col_types = col_brandon)   
dat2$Experience <- replace(dat2$Experience, dat2$Experience == "R", "0")
dat2$Experience <- as.integer(dat2$Experience)
#???ASK TA"nba2017-player-statistics.csv", col_types?????????????

str(dat2)
#read.csv - base
dat <- read.csv(file = "/Users/brandon/Desktop/stat133/stat133-hws-fall17/hw02/data/nba2017-player-statistics.csv", stringsAsFactors = FALSE, colClasses = colClasses)
dat$Experience <- replace(dat$Experience, dat$Experience == "R", "0")
dat$Experience <- as.integer(dat$Experience)
str(dat) 
``` 

### 4. Performance of Players
```{r}
#new statistics
Missed_FG = dat$FGA - dat$FGM
Missed_FT = dat$FTA - dat$FTM 
PTS = dat$Points3 * 3 + dat$Points2 * 2  + dat$FTM
REB = dat$OREB + dat$DREB
MPG = dat$MIN/ dat$GP


#add columns
dat <- cbind(dat, Missed_FG)
dat <- cbind(dat, Missed_FT)
dat <- cbind(dat, PTS)
dat <- cbind(dat, REB)
dat <- cbind(dat, MPG)

```

```{r}
EFF = ((dat$PTS + dat$REB + dat$AST + dat$STL + dat$BLK - dat$Missed_FG - dat$Missed_FT - dat$TO) / dat$GP)

summary(EFF) 

dat <- cbind(dat, EFF)

hist(dat$EFF, main = "Histogram of Efficiency", xlab = 'EFF', ylab = 'Frequency', ylim = c(0, 200))
```

``` {r}
#top then players by EFF
head(dat[order(dat$EFF, decreasing = TRUE), c(1, 2, 5, 30)], 10)

#players with negative EFF
dat[dat$EFF < 0, 1]
```

``` {r efficiency, fig.width = 12, fig.height = 8}
#correlation between EFF and variable 
co <- c(PTS = cor(dat$EFF, dat$PTS), 
REB = cor(dat$EFF, dat$REB),
AST = cor(dat$EFF, dat$AST),
STL = cor(dat$EFF, dat$STL),
BLK = cor(dat$EFF, dat$BLK),
Missed_FG = cor(dat$EFF, dat$Missed_FG) * -1,
Missed_FT = cor(dat$EFF, dat$Missed_FT) * -1,
TO = cor(dat$EFF, dat$TO) * -1,  
GP = cor(dat$EFF, dat$GP))

co <- sort(co, decreasing = TRUE) 

barplot(co, space = 0.3, col = ifelse(co < 0, 'red', 'grey'), ylim = c(-1,1))   
```

### 5. Efficiency and Salary
``` {r}
#plotting EFF (x) and Salary (y), include lowess smooth line, compute linear correlation coeffecient 

mil_salary <- dat$Salary/1000000

plot(dat$EFF, mil_salary, main = "Scatterplot of Efficiency and Salary", xlab = 'Efficiency', ylab = 'Salary (in millions)')

cor(dat$EFF, mil_salary)

lines(lowess(dat$EFF, mil_salary), col= 'red')
text(30, 30, 'lowess', col = 'red')
```

What can you say about the relationship between Efficiency and Salary? 
The relationship between Efficiency and Salary is that if efficiency goes up, salary is most likely to follow. The correlation is 0.655 which indicates they are positively correlated. 

``` {r}
#scatterplot for MPG > 20
players2 <- dat[dat$MPG > 20, ]
mil_salary_players2 <- players2$Salary/1000000

plot(players2$EFF, mil_salary_players2, main = "Scatterplot of Efficiency and Salary of Players with MPG > 20", xlab = 'Efficiency', ylab = 'Salary (in millions)')

lines(lowess(players2$EFF, mil_salary_players2), col= 'red')
text(32, 27, 'lowess', col = 'red')

cor(players2$EFF, mil_salary_players2)
```

What can you say about the relationship between these two variables for the "more established players"?
With MPG > 20 taken into account, the correlation is less than the relationship between efficency and salary of all players. However it is still positive which indicates a positive relationship between the two. If efficiency goes up, salary is likely to go up too. I think that there is a difference in correlation between the two because players with more experience's payrates do not fluctuate as much as the total with rookies. 

### 6. Comments and Reflections

Using read_csv was hard because a lot of the syntax changes and there were small things like col_factor you have to include the levels. I needed help on the col_factor and csv_read part. I did not know I had to manually put -1 into the Missed Field Goal and Missed Free Throw. The most time consuming part was read_csv. I am not sure if I fully understand the interpretation of the graphs for scatterplots of efficiency and salary of players. There wasn't any frustrating and figuring out the csv_read through a friend was the best part. 
